---
title: "Restaurant Sales Data Cleaning"
output: html_notebook
by : Juliatera Tambunan
date : 2025/08/17
---

# About Project

This is a Data Cleaning Project using Restaurant Sales Data from **Kaggle.** Download file [Here](https://www.kaggle.com/datasets/ahmedmohamed2003/restaurant-sales-dirty-data-for-cleaning-training "Download").

There are a couple steps in this cleaning process. Start from Here :

#### Load Library

I'm using Tidyverse package in this project 'cause i want to keep it simple and easy to read 😀

```{r}
library(tidyverse)
```

#### Import Data

Importing CSV file to RStudio

```{r}
sales_df <- read.csv("C:\\Users\\djoel\\Documents\\Data Analisis Bootcamp\\Project\\Data Cleaning\\R\\restaurant_sales_data.csv")
```

#### View NA Values

```{r}
colSums(is.na(sales_df))
```

#### Remove Outlier in Order.Total variable

Since there are many NA values in Order.Total, I want to exclude NA values so that this data can be analyzed further. Then create new data frame `sales_df_cleaned`

```{r}
sales_df_cleaned <-  sales_df %>% 
  drop_na(Order.Total)
```

#### Populate NA in Price with Order.Total/Quantity

There are NA values within Price variable, where all the Items should have price for each items. So i want to aggregate the NA values with `Order.Total / Quantity`.

```{r}
sales_df_cleaned <-  sales_df_cleaned %>% 
  mutate(Price = Order.Total / Quantity)
```

#### Check empty values in Item

Besides NA, there are more empty values in Item variable.

```{r}
sales_df_cleaned %>% 
  select(Item, Price) %>% 
  distinct() %>% 
  arrange(Price)
```

#### Populate empty value in Item based on Price

I want to populate Item that has empty value based on Price that corresponds to the Item.

```{r}
sales_df_cleaned <-  sales_df_cleaned %>% 
  select(Order.ID:Payment.Method) %>% 
  mutate(Item = if_else(Item == "" & Price == 2.5, "Iced Tea", Item),
         Item = if_else(Item == "" & Price == 15, "Grilled Chicken", Item),
         Item = if_else(Item == "" & Price == 1, "Water", Item),
         Item = if_else(Item == "" & Price == 3, "Lemonade", Item),
         Item = if_else(Item == "" & Price == 4, "Garlic Bread", Item),
         Item = if_else(Item == "" & Price == 5, "Ice Cream", Item),
         Item = if_else(Item == "" & Price == 6, "Brownie", Item),
         Item = if_else(Item == "" & Price == 7, "Cheesecake", Item),
         Item = if_else(Item == "" & Price == 8, "Chicken Melt", Item),
         Item = if_else(Item == "" & Price == 10, "Nachos Grande", Item),
         Item = if_else(Item == "" & Price == 12, "Pasta Alfredo", Item),
         Item = if_else(Item == "" & Price == 14, "Vegetarian Platter", Item),
         Item = if_else(Item == "" & Price == 18, "Salmon", Item),
         Item = if_else(Item == "" & Price == 20, "Steak", Item),
         )
```

#### Populate empty values in Payment.Method with "Unknown"

In `Payment.Method` there are empty value too. Since i lack of information of what the customers used to paid their orders, so i'll leave it with "Unknown".

```{r}
sales_df_cleaned <-  sales_df_cleaned %>% 
  mutate(Payment.Method = if_else(Payment.Method == "", "Unknown", Payment.Method))
```

#### Change Order.Date format to date

I realized the `Order.Date` variable had 'chr' class, so i changed it to 'Date'.

```{r}
sales_df_cleaned <-  sales_df_cleaned %>% 
  mutate(Order.Date = as.Date(Order.Date))
```

#### Check timeline

For additional information, i wanted to check the time range in this data, so i can see the trend line in further analysis.

```{r}
sales_df_cleaned %>% 
  select(Order.Date) %>% 
  summarise(MIN = min(Order.Date), MAX = max(Order.Date))
```

#### Export CSV

Finally i export the final data as cleaned Data in CSV format.

```{r}
write.csv(sales_df_cleaned, "C:\\Users\\djoel\\Documents\\Data Analisis Bootcamp\\Project\\Data Cleaning\\R\\restaurant_sales.csv", row.names = FALSE)
```

#### Final Result ( Before - After)

```{r}
View(sales_df) # Before
View(sales_df_cleaned) # After
```
